A Comparison of Dimensionality Reduction Techniques for Robotic Motion Planning

### 2016-05-13 paper/paper.tex
Merge conflicts to follow

The Rapidly exploring Random Tree algorithm works by building up a tree from
the initial configuration to the goal configuration \cite{lavalle1998rrt}.
Starting with the initial configuration as the root, it incrementally attempts
to add randomly selected samples to the tree until it reaches the goal
configuration.  The tree is ``rapidly exploring'' because the sample selection
is biased towards areas with large Voronoi regions (where the Voronoi regions
are computed based off of the nodes of the tree). RRTConnect, which is
described in Algs~\ref{alg:extend_RRT},~\ref{alg:connect_RRT},
and~\ref{alg:RRTConnect}, differs from RRT in that two trees, one with its root
at the initial configuration and one with its root at the start configuration,
grow towards each other \cite{kuffner2000rrt}.
%RRT*, which is described in Alg~\ref{alg:RRTstar} differs from RRT in that for every new node, connections are tested to the new node from every other node that is within a certain radius from the new node \cite{karaman2011sampling}. The the tree is pruned so that edges that are not part of the shortest path from the root of the tree to the new node are removed.

@@

Next we made a random obstacle generator. We create n-dimensional convex
polytopes by sampling $n^{2}$ random points on the n-sphere. We did this for
each point by choosing $n-1$ random angles and then defining $x_1$ through
$x_n$ using the equations below.

\begin{align*}
x_1 &= r \cos(\phi_1) \\
x_2 &= r \sin(\phi_1) \cos(\phi_2) \\
x_3 &= r \sin(\phi_1) \sin(\phi_2) \cos(\phi_3) \\
    &\vdots\\
x_{n-1} &= r \sin(\phi_1) \cdots \sin(\phi_{n-2}) \cos(\phi_{n-1}) \\
x_n &= r \sin(\phi_1) \cdots \sin(\phi_{n-2}) \sin(\phi_{n-1}) \,.
\end{align*}

After generating random obstacles, we ran Feature Agglomeration, Truncated SVD,
PCA, and Randomized PCA to reduce the dimensionality of the space. We then ran
PRM* and RRTConnect on each of these reduced dimensionality spaces, as well as
on the full-dimensionality space as a control. We then ran an inverse transform
on the paths that we find to return them to the full-dimensionality space. We
collected data on the number of runs that fail, the time it takes to run the
algorithm, and the path length returned. A summary of this paragraph can be
found in Alg~\ref{alg:DR}.

>>>>

The Rapidly exploring Random Tree algorithm works by building up a tree from
the initial configuration to the goal configuration \cite{lavalle1998rrt}.
Starting with the initial configuration as the root, it incrementally attempts
to add randomly selected samples to the tree until it reaches the goal
configuration.  The tree is ``rapidly exploring'' because the sample selection
is biased towards areas with large Voronoi regions (where the Voronoi regions
are computed based off of the nodes of the tree). RRTConnect, which is
described in Algos.~\ref{alg:extend_RRT},~\ref{alg:connect_RRT},
and~\ref{alg:RRTConnect}, differs from RRT in that two trees, one with its root
at the initial configuration and one with its root at the start configuration,
grow towards each other \cite{kuffner2000rrt}.
%RRT*, which is described in Alg~\ref{alg:RRTstar} differs from RRT in that for every new node, connections are tested to the new node from every other node that is within a certain radius from the new node \cite{karaman2011sampling}. The the tree is pruned so that edges that are not part of the shortest path from the root of the tree to the new node are removed.

@@

Next we made a random obstacle generator. We create n-dimensional convex
polytopes by sampling $n^{2}$ random points within the $n$-sphere. To generate
a single point, you must sample $n - 1$ angles from a uniform distribution
between 0 and $2\pi$, $n$ radii from a Gaussian distribution,
$\mathcal{N}(r_{\mu}, r_{\sigma})$ where $r_{\mu}$ and $r_{\sigma}$ are the
mean and standard deviation for the radius of the convex polytope, and $n$
floats to act as the center of the $n$-sphere.  Then you use the series of
equations in Eq.~\ref{eq:nsphere} to produce a randomly sampled point within
the $n$-sphere where $c_k$, $r_k$, and $\phi_k$ are the centers, radii, and
angles respectively.  To produce a convex polytope, we then computed the
$n$-dimensional convex hull over these points. The convex hull then provided
the normal equations and offsets for each facet on the polytope.

\begin{align}
    \begin{split}
        x_1 &= c_1 + r_1 \cos(\phi_1) \\
        x_2 &= c_2 + r_2 \sin(\phi_1) \cos(\phi_2) \\
        x_3 &= c_3 + r_3 \sin(\phi_1) \sin(\phi_2) \cos(\phi_3) \\
            &\vdots\\
        x_{n-1} &= c_{n - 1} + r_{n - 1} \sin(\phi_1) \cdots \sin(\phi_{n-2}) \cos(\phi_{n-1}) \\
        x_n &= c_n + r_{n} \sin(\phi_1) \cdots \sin(\phi_{n-2}) \sin(\phi_{n-1}) \,.
    \end{split}
    \label{eq:nsphere}
\end{align}

After generating random obstacles, we ran Feature Agglomeration, Truncated SVD,
PCA, and Randomized PCA to reduce the dimensionality of the space by using the
sampled points within each $n$-sphere as training data. We then ran PRM* and
RRTConnect on each of these reduced dimensionality spaces, as well as on the
fully-dimensional space as a control. We then ran an inverse transform on the
paths that we find to return them to the fully-dimensional space in order to
verify that the path in the reduced dimension did not come into collision with
obstacles. We collected data on the number of runs that fail to find a
solution, the time it takes to run the algorithm, and the length of the
solution if one was found.  A summary of this paragraph can be found in
Algo.~\ref{alg:DR}.

For our experiments, we varied the number of random obstacles in the scene, the
planning algorithm used, and the method of dimensionality reduction. For each
problem instance, the random generator was seeded in order to make fair
comparisons between the planners and dimensionality reduction techniques.
Also, since the algorithms are non-deterministic, each problem instance
executed 100 times. We used scenes with 250, 300, and 350 obstacles in order to
measure how the planners and dimensionality reduction techniques vary in their
performance with varying levels of obstacle density. In the results presented,
we are reducing from 8 dimensions to 6 dimensions and using mean radius of
$14.8$ for the convex polytopes.
